---
title: "Sara_Herrera_Q15"
author: "Sara Herrera (PID:A59011948)"
date: "7/13/2022"
output:
  pdf_document: default
  html_document: default
---

# Name: Sara Herrera de la Mata
# PID: A59011948

# Covid-19 Variants in California

**Libraries**

```{r}
# Calling the libraries that will be used to generate the plot
library(ggplot2) # For plots
library(lubridate) # For date-time data
library(dplyr) # For data manipulation
```

**Read file**

```{r}
# To read the .csv file downloaded from the California Health and Human Services (CHHS) open data site
data <- read.csv("covid19_variants.csv")
```

**Filter data**

I'll use 'dplyr' to filter out unnecessary rows (containing the values 'Other' and 'Total' in the 'variant_name' column) and change the format of the date values in the 'date' column to date instead of character.

```{r}
# To remove the rows that contain "variant_name" = "Other", "Total" (as they're not included in the plot)
# To change the date values from "character" to "date" format
clean.data <- data %>% filter(!variant_name %in% c('Other', 'Total')) %>% mutate(date = as.Date(date))
```

**Assigning variables**

We use this to simplify things when writing code, substituting longer names to simpler and shorter words.

```{r}
# Assigning variables
Percentage <- clean.data$percentage
Date <- clean.data$date
Variant <- clean.data$variant_name
```

**Plotting the data** 

I'll use 'ggplot2' and I'll first specify the data frame and the variables that we're interested in, coloring each Covid-19 variant with a different color (each variant will be a line in the graph). Then, I'll select the type of graph I would like to represent the data in, in this case I'll use a line chart. I'll modify the theme of the plot so that it resembles the example and I'll modify the dates with 'lubricate'. In addition, I'll add labels to my plot so that it's easily understandable and a caption at the bottom.

```{r}
# Making the plot
plot <- ggplot(data = clean.data, aes(x=Date, y=Percentage, color=Variant)) +
  # Select the data frame, x and y axis variables, and the variable by which the data will be colored by
  geom_line(aes(group=Variant)) +
  # To make a line chart, where every line will be a category of the "Variant" column, meaning a different Covid-19 variant
  theme_bw() +
  # Specifies the theme used for the plot: in this case a white background and thin grey grid lines
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  # Specifies how the dates in the x-axis will be represented, separated by 1 month gaps, and with the label month (in letters abbreviated) and year
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  # Specifies that the label for the dates in the x-axis is angled
  labs(title = "Covid-19 Variants in California", x = NULL, color = NULL, caption = "Data Source: <https://www.cdph.ca.gov/>") +
  # To label the title of the plot, remove the x-axis label to only keep the dates, remove the label of the variants to only keep the name of each variant next to each associated color, and to add a caption at the bottom of the plot
  ylab("Percentage of sequenced specimens") # To add a y-axis label 
plot
```
