---
title: "class17.Rmd"
author: "Sara Herrera (PID:A59011948)"
date: "11/24/2021"
output: 
  pdf_document:
    latex_engine: xelatex
---

# COVID-19 Vaccination Rates

```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```
> How many entries do we have?

```{r}
nrow(vax)
```

We can use the **skimr** package and the `skim()` function to get a quick overview of the structure of the dataset:

```{r}
skimr::skim(vax)
```

We noticed that one of these columns is a date column. Working with time and dates get's annoying quickly, and there's a package to help with this called **lubridate**:

```{r}
library(lubridate)
today()
```

> How many days since the first entry in the dataset?

```{r}
vax$as_of_date[1]
```
We need to change the data column from character to numerical:
```{r}
# today() - vax$as_of_date[1] won't work

d <- ymd(vax$as_of_date)
d[1]
today() - d[1]
```

I will make the `as_of_date` column Date format:

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

> When was the dataset last updated? What is the last date in this dataset? OR How many days since the last update?

```{r}
today() - vax$as_of_date[ nrow(vax) ]
```

> How many days does the dataset span?

```{r}
vax$as_of_date[ nrow(vax) ] - vax$as_of_date[1]
```

## Working with ZIP codes

> How many different Zip code areas are recorded in this dataset?

```{r}
length(unique(vax$zip_code_tabulation_area))
```

To work with Zip codes we can use the **zipcodeR** package:

```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

## Focus on the San Diego area

We want to subset the full CA `vax` data down to just San Diego county.
We can do it with base R:

```{r}
# Subset to San Diego county only areas
inds <- vax$county == "San Diego"
nrow(vax[inds,])
```

But much better with the **dplyr** package to make things easier and faster and be able to subset by:

```{r}
library(dplyr)

# Use `filter()` function to do our subsetting:
sd <- filter(vax, county == "San Diego")
nrow(sd)
```
More complicated filtering:
```{r}
sd.20 <- filter(vax, county == "San Diego" &
                age5_plus_population > 20000)
nrow(sd.20)
```

> What is the average vaccination rate of San Diego county as of the last updated date?

```{r}
sd.now <- filter(vax, county == "San Diego" &
                as_of_date == "2021-11-23")
head(sd.now)
```

```{r}
# `summary` function is a good way to get mean, median... fast
summary(sd.now$percent_of_population_fully_vaccinated)
```

> Make a histogram of sd.now

Base R histogram:
```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```

This is going to be very susceptible to be skewed to small population zip codes, where vaccinated or unvaccinated people will have a very big effect on small numbers.

## Focus on UCSD/La Jolla

> What is the population of the 92037 Zip code?

Using Zip code 92037
```{r}
lj <- filter(sd.now, zip_code_tabulation_area=="92037")
lj$age5_plus_population
```

> What is the average vaccination for this Zip code?

```{r}
lj$percent_of_population_fully_vaccinated
```

> What about this Zip code 92122?

```{r}
lj2 <- filter(sd.now, zip_code_tabulation_area=="92122")
lj2$age5_plus_population
```
```{r}
lj2$percent_of_population_fully_vaccinated
```

> What about 91942?

```{r}
lj3 <- filter(sd.now, zip_code_tabulation_area=="91942")
lj3$age5_plus_population
```
```{r}
lj3$percent_of_population_fully_vaccinated
```

We'll do a time series of vaccination rate for a given Zip code area: La Jolla
Start with 92037.

```{r}
# We need to go back to full vax dataset, as the sd.now only has yesterday's date on it.

lajolla <- filter(vax, zip_code_tabulation_area=="92037")

library(ggplot2)

ggplot(lajolla) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```

## Comparing to similar sized areas

Let's make this plot for all San Diego county areas with populations at lease as large as the Zip code 92037.

```{r}
sd.36 <- filter(vax, county=="San Diego", age5_plus_population > 36144)
head(sd.36)
```

> How many Zip code areas in SD county have a population bigger than 92037?

```{r}
length(unique(sd.36$zip_code_tabulation_area))
```

Let's make this plot with all the Zip codes:

```{r}
ggplot(sd.36) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated,
      group=zip_code_tabulation_area) +
  geom_line() +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```

> Make a plot for the whole state of CA for all Zip code areas with population at least as large as La Jolla

```{r}
ca <- filter(vax, age5_plus_population > 36144)
```

How many Zip codes are we talking about?

```{r}
length(unique(ca$zip_code_tabulation_area))
```

```{r}
ggplot(ca) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated,
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated") +
  geom_hline(yintercept = 0.67, color="red")
```

> What is the mean across the state for these 36k + population areas?

```{r}
# The mean line was applied in the plot above.
ca.now <- filter(ca, as_of_date=="2021-11-23")
summary(ca.now$percent_of_population_fully_vaccinated)
```

## About this document

```{r}
sessionInfo()
```




















